\documentclass{article}
\usepackage{graphicx} % Required for inserting images and \rotatebox
 \usepackage[tight]{subfigure}
 \usepackage{tikz}
 \usetikzlibrary{shapes, arrows, positioning, fit, backgrounds}
 \usepackage{amssymb}
 \usepackage{amsmath}
 \usepackage{clrscode3e}
 \usepackage{tablefootnote}
 \usepackage{hyperref}
 \usepackage{xcolor,colortbl} % for row colors and cell colors
 \usepackage{booktabs}  % for \toprule, \midrule, \bottomrule
\usepackage{pifont}    % for \xmark alternative
\newcommand{\xmark}{\ding{55}} % for table
\newcommand{\qmark}{?} % for table

 
 \usepackage{natbib}
\setlength{\bibsep}{0.5pt}

\setlength{\oddsidemargin} {0 in}
\setlength{\textwidth} {6.5 in}
\setlength{\textheight} {8.5 in}
\setlength{\topmargin} {-0.5 in}
\def\h{\hspace{-.9pt}{\_}}
\newcommand{\reminder}[1]{ [[[ \marginpar{\mbox{$<==$}} #1 ]]] }
\newcommand{\eatreminders}[0]{\renewcommand{\reminder}[1]{}}

\title{Privacy Preserving Compliance}
\author{Joss Duff}
\date{\today}

\begin{document}
\eatreminders

\maketitle

\reminder{Joss: General and consistency TODOs:
 \begin{itemize}
     \item Capitalization on section titles
     \item Zk abbreviations and hyphenations
     \item privacy-preserving vs privacy preserving
     \item Merkle capitalization
     \item Implementation code link somewhere
 \end{itemize}
}

\section{Abstract}

Privacy and compliance in blockchain systems face an inherit contradiction: privacy requires that only the individual knows certain information, while compliance traditionally requires revealing information to prove compliance criteria.  This tension has led to two unsatisfactory outcomes:
\begin{enumerate}
    \item Privacy systems without compliance attract illicit actors and face legal challenges.
    \item Compliance systems without privacy expose all transaction histories publicly, making them unacceptable for consumer and commercial adoption.
\end{enumerate}
Neither outcome supports the widespread adoption of blockchain technology, particularly among traditional financial institutions where regulatory uncertainty and privacy concerns are the top barriers to entry.

We present a composable framework that resolves this conflict by enabling users to prove compliance cryptographically without revealing private transaction data.  In our system, regulatory bodies publish compliance definitions as combinations of reusable logical primitives called constraints.  Applications select relevant compliance definitions based on their regulatory obligations.  Users generate zero-knowledge proofs demonstrating that they meet these requirements, which are then verified on-chain before transactions execute.  This provides proactive compliance: preventing non-compliant actions rather than detecting them afterward, all while maintaining strong privacy guarantees.

Our framework is distinguished by its composability and generality, designed explicitly for a global ecosystem with many regulatory bodies whose requirements evolve over time.  Compliance definitions are chain-agnostic, modular with respect to privacy mechanisms, versioned to accommodate regulatory updates, and can be combined arbitrarily which allows applications to satisfy multiple jurisdictions simultaneously.


\section{Background and Definitions}
Background information on blockchains, privacy systems on blockchains, ZK proofs, and our working definition of ``compliance".  People familiar with blockchain and ZK can skip this section.

\subsection{Zero Knowledge Proof}
\label{sec:zk}
Zero Knowledge (ZK) proofs are a cryptographic technique that allows a party to prove specific claims about a computation without revealing any additional information about the computation.  The origins of this concept
predate blockchain~\cite{ZK88,ZK-origins}, but it was described elegantly and simply by Vitalik Buterin et al. in~\cite{buterin2023blockchain}
as follows:
\begin{quote}
The ``claim'' proven by a ZK proof is expressed as a type of program that is often called a \textit{circuit}.  Mathematically, it suffices to think of it as a function $f(x, w) \rightarrow \{\text{True}, \text{False}\}$, where $x$ is the public input, $w$ is the private input, and $f(\cdot)$ is the function being computed. A ZK proof proves that, for a given $x$ known both by the prover and the verifier, the prover knows a $w$ such that $f(x, w)$ returns $\text{True}$.  
\end{quote}


SNARKs are a specific type of ZK proof that has the unique property of \textit{succinctness}: The proof size doesn't grow with the size of the computation, and the proof can be verified very quickly relative to the computation~\cite{Petkus}.\footnote{The SNARK acronym comes from Succinct, Non-interactive ARguments of Knowledge. The non-interactive term serves to distinguish SNARKs from the interactive proofs between a prover and a verifier that are often used to illustrate the concept of ZK for a general audience.}

A drawback with some SNARK schemes, such as Groth16~\cite{Groth:2016}, is that they rely on a \textit{trusted setup}: a time-intensive ceremony involving multiple participants.  A trusted setup needs to be performed once per circuit~\cite{lavin2024surveyapplicationszeroknowledgeproofs} for Groth16 and some other SNARKs.

A type of SNARK proofs, PLONK, mitigates this drawback by allowing trusted setups to be re-used across circuits.  With PLONK proofs, it is not necessary to initiate a new trusted setup for every new circuit~\cite{Gabizon2019PLONKPO}.

\subsection{Compliance}
\label{sec:compliance}
\textit{Compliance} is referred to as the goals and properties of a system to attempt to adhere to a law or policy.  Compliance isn't immunity from prosecution, it is a best effort to prove legality of actions.  For example, US companies have systems in place to prevent money laundering in order to be compliant with the US Bank Secrecy Act (BSA) - also referred to as Anti-Money Laundering (AML)~\cite{usc31_5318h}.  A typical effort to comply with AML in the blockchain space is blocking addresses associated with sanctioned individuals from interacting with a protocol.  

A goal of this framework is to be flexible enough to accommodate compliance definitions over all aspects of a transaction: the payer, the payee, the content of the transaction, and even the actors' transaction histories.  Flexibility results in the ability to express any definition of compliance.  This framework is built around the fact that definitions of compliance vary over time and by jurisdiction.


\subsection{Blockchain Background}
Blockchains are fundamentally data-storage systems with unique properties.  Most importantly, data stored on a blockchain (on-chain) are \rm{immutable}: history cannot be changed.  This property is achieved by allowing anyone  around the world to participate in consensus and validation of the blockchain's state by running a node.  There is a small class of \textit{permissioned} blockchains with access restrictions on who can use the chain or run a node.

Nodes have a financial incentive to participate correctly in consensus.  This mechanism makes it economically infeasible for a single attacker to disrupt the network.  For example, in the Ethereum network, nodes who want to participate in block production must put up a stake of the Ether currency\footnote{The original staking requirement was 32 Ether but it is subject to change.} which is \textit{``slashed"} (lost forever) if the node behaves incorrectly\cite{eip2982}.

The first blockchain, Bitcoin, was created in 2008 and proposed a trustless ledger of transaction histories\cite{Bitcoin}.  Ethereum was created in 2013 and improved on Bitcoin by allowing trustless execution of Turing-complete logic on-chain, referred to as \textit{smart-contracts}~\cite{Buterin13}.  This allowed for more complex applications like marketplaces and lending platforms to exist in a decentralized, permissionless, and immutable setting. 

Every transaction on a blockchain has a non-static monetary cost paid to the block-producing node as an incentive to include the transaction in the next write.  Some blockchains with smart-contract support, like Ethereum, include an additional transaction cost driven by the complexity of the smart-contract logic being executed.  

Users on blockchains are represented by a public-key address.  The user signs messages and transactions with the associated private key.  Users can have multiple public-key addresses.  It is impossible to reveal who is associated with a public key unless the users identity is revealed.  Identity can be revealed directly by the user or indirectly by others who know the public key's identity, like a centralized exchange.

A centralized exchange (CEX) is a company that allows users to trade traditional currency for cryptocurrency.  They often require the user to connect their bank account and supply regulatory documentation such as identification cards and proof of residency.  This is a common method for acquiring cryptocurrency.  CEXs are \textit{custodial}: they control the address (they know the private key) that owns the cryptocurrency on-chain and allow the user the ability to trade or transfer the cryptocurrency.  They also allow users to send their cryptocurrency from the CEX account to the users' own on-chain address.  CEX accounts are not addresses.  Rather, the CEX keeps its own addresses and the user's holdings are only an accounting record in the CEX's books.

A decentralized exchange (DEX) enables trading cryptocurrency while letting users retain direct control of their holdings.  In its simplest form, the DEX creates a smart-contract to implement the trade and the users sign it with their private key.

\subsection{Privacy in the Blockchain Setting}
All data on blockchains are public by default unless the chain supports special privacy features; Aztec Network~\cite{aztecprotocol}, Zcash~\cite{ZcashWhitepaper}, and others.  There are also some enterprise permissioned blockchains that restrict visibility levels.  However, the vast majority of chains are public by default, including Ethereum and Bitcoin.

For public blockchains, every user's entire transaction history is exposed to the world.  This is obviously undesirable for consumers.  Most people don't want their purchase or subscription history visible to the entire world.  Many applications and protocols exist to provide users varying degrees of privacy.

Blockchain privacy protocols fall into two broad categories: 
\begin{itemize}
    \item \textit{App-layer privacy}: Applications on top of existing blockchains.
    \item \textit{Base-layer privacy}: Entirely new blockchain architectures with privacy as a core identity.
\end{itemize}

\subsubsection{App-layer privacy}
CoinJoin was the first permissionless app-layer transaction mixer~\cite{maxwell2013coinjoin}.  CoinJoin was built on Bitcoin and combined multiple input UTXOs from different parties into multiple outputs, obscuring which inputs correspond to which outputs.  Tornado Cash was the first transaction mixer to use ZK proofs~\cite{tornadocash}, and saw massive popularity~\cite{defillama2025tornado}.  ZK proofs are a common tool used to achieve privacy; ZK allows a party to prove a computation without revealing the entire computation.  See section \ref{sec:zk} for an introduction to ZK proofs.  The black-hat group Lazarus used Tornado Cash to aid in laundering stolen funds and the Tornado Cash developers were put on trial~\cite{doj_tornado_cash_2023}.  Railgun~\cite{railgun} and Hinkal~\cite{hinkal} are transaction mixers that focuses on compliance.  Privacy Pools is a recent transaction mixer that also focuses on compliance with \textit{association sets}~\cite{buterin2023blockchain}.  Ultra-Anon is an experimental privacy token that has the property of \textit{plausible deniability}: it isn't revealed if users utilized privacy~\cite{ultraanon}.  Stealth addresses are ephemeral public keys that can receive any tokens while hiding the receiver address~\cite{buterin2023stealth}.  zkBob is a wallet that has private transfer for certain stablecoins~\cite{zkbob2023overview}.  Webb Protocol creates a cross-chain anonymity set~\cite{webbprotocol}.  Worm is an application implementation of ``EIP-7503: Zero-Knowledge Wormholes"~\cite{7503}, which is another transactional privacy solution with the property of plausible deniability.

\subsubsection{Base-layer privacy}
Zcash is a privacy blockchain based on Bitcoin's \textit{unspent transaction output} (UTXO) model.  Zcash is implemented from the 2014 Zerocash paper~\cite{ZcashWhitepaper}.  Monero is another private blockchain that leverages the UTXO system and is based on the 2013 CryptoNote paper~\cite{cryptonote}.  Grin is an implementation of the 2016 Mimblewimble protocol, a modification of Bitcoin with private transfers~\cite{jedusor2016mimblewimble}.  Aztec is a general purpose blockchain that allows for private state with smart-contract support~\cite{aztecprotocol}.  Aztec is an Ethereum layer 2.  Aleo~\cite{aleo2025homepage}, based on the 2018 paper ``ZEXE: Enabling Decentralized Private Computation"~\cite{ZEXE}, is also a blockchain with private state and smart-contracts.  The Miden blockchain supports the private state similarly~\cite{miden2024web}.  Penumbra is a private network and decentralized exchange for the Cosmos ecosystem~\cite{penumbraprotocol}.  Arc~\cite{circlearc} and Tempo~\cite{tempo2025}, founded by Circle and Stripe, respectively, are stablecoin-focused blockchains with some degree of transaction privacy.  Payy is an Ethereum layer 2 with native transaction privacy and compliance mechanisms~\cite{payywhitepaper}.  Secret Network has private by default smart contract state~\cite{secretnetwork2025docs}.  Ernst \& Young (EY) released a permissioned layer 2 blockchain, Nightfall, with private transactions and compliance features~\cite{ey2025blockchain}.

\subsubsection{Custodial mixers}
\textit{Custodial mixers} are transaction privacy solutions that require users to send funds to a trusted intermediary who then redistributes the funds.  This accomplishes the privacy goal of breaking the on-chain traceability between the input and output transactions, but custodial mixers have the major weakness that operators could steal funds or keep records.  This is fundamentally weaker privacy because that intermediary has knowledge of a user's transaction history, rather than that history being private and held only by the user.  While this is not a legal document, we note that in some jurisdictions, the intermediary could be subject to government subpoena.  

Blender.io was an example of a custodial mixer on Bitcoin where users sent funds to a trusted operator who then re-distributed the funds.  Blender was sanctioned in 2022 by the U.S. Treasury Department for failing to meet AML regulations~\cite{sun2022blender}.

Some centralized exchanges can also be used as custodial mixers.  A user deposits into the centralized exchange from one address and withdraws from another address.  From the on-chain perspective, this looks like two unrelated transactions among the millions that occur within that centralized exchange. However, the centralized exchange has full information and knows the two addresses are owned by the same user.

\subsubsection{Anonymity Set}
An \textit{anonymity set} is the group of entities among which an observer cannot reliably identify which specific entity possesses a particular attribute or performed a particular action~\cite{anonterminology,anonset}.  Anonymity scales with set size: the larger the anonymity set, the less identifiable each entity is~\cite{Dingledine2006AnonymityLC}.

For transaction mixer applications on blockchains like Railgun or Tornado Cash, any address that has deposited can withdraw, but it is never revealed which depositing addresses have an associated withdrawal.  Each withdrawal could have come from any of the depositing addresses.  Therefore, the anonymity set of any given withdrawal from the mixer is the set of all addresses that have deposited in the mixer.  This is oversimplified, and in practice anonymity sets can be reduced with chain analysis~\cite{zcashanalysis,moneroanalysis}.  Anonymity here is referring to the obfuscation of the link between a depositing and withdrawing address.

\section{Problem} 
Blockchains are fundamentally pseudonymous and public: user identities are obfuscated behind addresses, but all activity between addresses is public by default.  In practice, entities are able to associate identities to addresses when on-ramping traditional funds to cryptocurrency and therefore track individuals' flows of funds on-chain.  In order for cryptocurrency to be practical for consumers, it must at the very least have the same level of privacy as traditional finance, but ideally improves on traditional finance privacy.  In traditional finance, typically two entities know information about a user's funds; the user and the exchange that is custodian of the funds.  In blockchain settings, all entities can track user funds and activity.  Neither consumers nor corporations will want their entire trading, purchasing, or subscription histories publicly available.

Privacy is a growing field in the blockchain space~\cite{tornadocash, 7503, 6956581, ultraanon}, but raises the concern that malicious actors may use privacy to launder funds.  In an effort to reduce money laundering, regulatory bodies are pursuing legal action against developers of privacy protocols~\cite{kuhn_samourai_2024,doj_tornado_cash_2023,bitcoin_fog_2021}.  This introduces the need to explicitly meet regulatory compliance inside private on-chain applications.

The goal is to maximize privacy while maintaining the ability to prove compliance.  Current research efforts attempting to bridge privacy and compliance have proposed solutions with strict domain limitations such as a new blockchain architecture, specific on-chain protocols, or privacy violations.  We investigate a framework that allows for compliance in any blockchain smart-contract environment and in any privacy protocol without compromising privacy provided by the protocol.

\subsection{Industry Interest}
The two privacy protocols currently operational that have some measure of compliance are Privacy Pools and Railgun.  Both have received a great deal of interest from users; Railgun was launched in 2022 and reached \$132 Million (USD) deposited~\cite{defillama2025railgun} in August 2025. Privacy Pools launched in March of 2025 and reached \$3.3 Million (USD) deposited~\cite{defillama2025privacypools} in January 2026.  \reminder{TODO update these figures in the final thesis}  Many users currently in the blockchain space desire both compliance and privacy.

Large traditional financial corporations are yet to integrate significantly with blockchains.  In a 2025 survey, 300 major financial firms were asked what factors keep them from becoming more involved in blockchain technology. In this multiple-choice question, 52\% cited lack of regulatory certainty and 50\% cited lack of privacy as factors~\cite{paradigm2025tradfi}.  These are two of the top three blocking factors.  We see it is not just users, but also institutions who desire compliance and privacy.  Solving this would result in a significant amount of capital being allocated on-chain.

In 2024, \$2.2 billion (USD) was stolen in crypto hacks and exploits~\cite{trmlabs2025cryptocrime}.  Tracing stolen funds from all hacks is too large a task, so auditors prioritize only the largest exploits.  Our framework provides strong \textbf{proactive} compliance guarantees.  Non-compliant actors are stopped from interaction with compliant applications.  Compare this to retroactive compliance measures where non-compliant actors are able to interact with private protocols freely, but there exists a key that is able do decrypt any user's transaction history to be used by law enforcement.  This retroactive approach is sometimes referred to as view-key compliance.  Not only does this create more work for auditors trying to trace non-compliant individuals, it also introduces the risk of leaking compliant user's private information.

\reminder{HFK: Here we can go beyond financial stuff. Supply chain is a good example here. It can help a B2B supply chain run on a public L1 underpinning.  Also within a permissioned system, it can reduce the degree of central control by the permissioning agent.  The concept could fit in the Oracle framework -- or at least Mark Rakhmilevich saw some merit to it. }

\section{Desired privacy preserving compliance properties}
\label{sec:privacy-preserving-compliance}
The definitions provide a path for a strong and flexible compliance solution without compromising the benefits of having a privacy system on-chain.

\paragraph{No deanonymization:}  \textit{deanonymization} refers to revealing private information about an address, like balances or transaction history.  No entity should have the ability to deanonymize user data in an on-chain privacy system\footnote{Caveat: A user can always choose to reveal their own private transaction history.  This is acceptable because the user is initiating their own deanonymization.  ``No deanonymization" refers to third-party deanonymization.}.

Any compliant privacy solution that gives privileges to decrypt private user data has deanonymization.  The ability for a third party to reveal a user's private details is fundamentally weaker privacy.  We can guarantee that private transactions can only be revealed by the user, and we provide this guarantee without compromising compliance.  We also provide stronger compliance guarantees than view-key compliance.  Viewing keys are used for retroactive compliance checks, and our framework has proactive compliance: non-compliant actors are stopped before making their transaction rather than after.  Additionally, with view-key compliance a user can never verify if their secrets have been read or shared.  In our solution a user's secrets are always kept private.

The existence of the ability to deanonymize users is a risk vector to users and compromises trustlessness, as highlighted in~\cite{a16z_zkp_privacy_2022}.  Individuals with access to private user information can and will abuse it.  Examples include the case of Twitter employees selling personal user data to Saudi Arabia~\cite{newman2019twitter, collier2022twitter}, or US telecom providers illegally selling user data~\cite{daws2024fcc}.  Additionally, any centralized collection of user data can be hacked.  In 2024, data that US telecommunication companies were legally required to collect was hacked~\cite{sakellariadis2024china}.  In 2025, personal data held by the Ukrainian government was hacked by Russia~\cite{post2025ukraine}.  In 2022, 23 terabytes of personal information on 1 billion Chinese residents was leaked from a police database and sold~\cite{wikipedia2025shanghai}.

On-chain privacy is unique because it allows for truly private-to-the-world transactions.  An on-chain privacy system where a third party also knows secret user data is as ``private" as a traditional exchange or bank.  In order for a on-chain privacy system to provide unique value to users, it must not allow for deanonymization.

\paragraph{Proactive compliance:} A compliance system is \textit{proactive} if it prevents non-compliant action rather than detecting it afterwards. A proactive compliance system both preserves user privacy by eliminating the need for information disclosure and reduces regulatory burden by automating compliance checks.  Regulators can check compliance by verification of proofs, which is much more efficient than having to evaluate full execution histories.  Users can show transparently that they are in compliance without having to reveal private data.  Our framework prevents non-compliant actions by requiring a proof of compliance before a transaction is executed.  

\paragraph{Rich compliance language:} Compliance can be expressed in many languages.  At the lowest level, it is the language of legal documents.  Expressing compliance on-chain is more restrictive because of the inherent setting.  Many protocols express compliance on-chain as only a set of sanctioned addresses.  We allow for the maximum expression of compliance by allowing any on-chain data to be used in a compliance definition: transaction histories, contract states, fund origins, user secrets, etc.  A rich compliance language allows for regulators to clearly define exactly what actions are non-compliant, leading to less ambiguity and requiring the regulators to decide on a concrete definition.

Additionally, we include time as a first class parameter compliance definitions.  This captures the invariant of real-world regulations changing over time.  The time parameters of compliance definitions are accessible on-chain: users are able to prove that they were compliant with a regulation at a previous time and the history of a compliance definition is transparent.

\paragraph{Multiple compliance:} There are many government bodies in the world, each with their own laws and regulations.  A suitable compliance protocol should be able to handle any amount of compliance definitions.  A compliance protocol that only supports one government's regulations is not suitable for an international finance system.

\paragraph{Secret inputs:} We allow for arguments in the compliance evaluation that are only known to the user.  This is trivial in most solutions that utilize ZK proofs for compliance verification, as ZK proofs inherently allow secret inputs.  However this is difficult to achieve in systems where the compliance evaluation is done on a remote server.

\paragraph{Verifiable compliance:} Compliance criteria should be publicly viewable and a user's compliance status should be verifiable.  A user's compliance status should not be determined behind a third-party API as this runs the risk of the third-party not properly enforcing the compliance definition, introducing risk to all parties involved.  This trust assumption could potentially be mitigated with techniques such as \textit{trusted execution environments} (TEEs) that verify that a third-party correctly executed some computation~\cite{haventee}.  The approach in this framework is to publish compliance definitions publicly and have users generate their own ZK proofs that can be cryptographically verified.

\paragraph{Chain agnostic:} A compliant privacy solution is \textit{chain agnostic} if it does not rely on any protocol level mechanism of specific blockchain.  Application layer solutions are chain agnostic by default because they exist on-chain as a set of smart contracts.

\paragraph{Modular privacy:} Some compliant privacy solutions are dependent on a specific implementation of on-chain privacy.  Privacy preserving compliance shouldn't depend on privacy mechanisms, as new privacy mechanisms are constantly being researched.  A compliant privacy solution has \textit{modular privacy} if it can be applied to any privacy protocol regardless of its mechanism.

\section{Related Work and Attempted Solutions}
\label{sec:related-work}
We review other work that attempts to solve the compliant privacy problem by a feature comparison.

\begin{table}[htbp]
\centering
\caption{Comparison of privacy preserving compliance solutions}
\label{tab:comparison}
\small
\rowcolors{2}{gray!15}{white}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
\textbf{Solution} & 
\rotatebox{90}{\textbf{No deanonymization}} & 
\rotatebox{90}{\textbf{Proactive compliance}} & 
\rotatebox{90}{\textbf{Rich compliance language}} & 
\rotatebox{90}{\textbf{Multiple compliance}} & 
\rotatebox{90}{\textbf{Secret inputs}} & 
\rotatebox{90}{\textbf{Verifiable compliance}} & 
\rotatebox{90}{\textbf{Chain agnostic}} & 
\rotatebox{90}{\textbf{Modular privacy}} \\
\hline
Aleo~\cite{aleo2025homepage} (ZEXE~\cite{ZEXE}) &\checkmark &\xmark &\xmark &\checkmark &\qmark &\xmark &\xmark &\checkmark \\
\hline
Payy~\cite{payywhitepaper} &\xmark &\xmark &\xmark &\xmark &\qmark &\xmark &\xmark &\checkmark \\
\hline
Arc~\cite{circlearc} &\xmark &\xmark &\xmark &\xmark &\xmark &\xmark &\xmark &\checkmark \\
\hline
Tempo~\cite{tempo2025} &\xmark &\qmark &\qmark &\checkmark &\xmark &\qmark &\xmark &\checkmark \\
\hline
ML Reg\tablefootnote{``A privacy-preserving scheme with multi-level regulation compliance for blockchain"~\cite{privacyregulation}}~\cite{privacyregulation} &\xmark &\xmark &\xmark &\xmark &\xmark &\xmark &\xmark &\checkmark \\
\hline
REGKYC~\cite{Xiong2025REGKYCSP} &\xmark &\qmark &\xmark &\checkmark &\checkmark &\qmark &\checkmark &\xmark \\
\hline
zkFi~\cite{chaudhary2025zkfiprivacypreservingregulationcompliant} &\xmark &\xmark &\xmark &\xmark &\xmark &\xmark &\checkmark &\xmark \\
\hline
Daze~\cite{hazedaze} &\xmark &\qmark &\xmark &\xmark &\xmark &\qmark &\checkmark &\xmark \\
\hline
Haze~\cite{hazedaze} &\checkmark &\checkmark &\xmark &\xmark &\xmark &\checkmark &\checkmark &\xmark \\
\hline
ZEBRA~\cite{Rathee2022ZEBRAAC} &\xmark &\xmark &\xmark &\xmark &\xmark &\xmark &\checkmark &\checkmark \\
\hline
SeDe~\cite{chaudhary2025sedebalancingblockchainprivacy} &\xmark &\xmark &\xmark &\xmark &\xmark &\xmark &\checkmark &\qmark \\
\hline
CP HABE\tablefootnote{``A Regulation Scheme Based on the Ciphertext-Policy Hierarchical Attribute-Based Encryption in Bitcoin System"~\cite{8314106}}~\cite{8314106} &\xmark &\qmark &\xmark &\xmark &\xmark &\checkmark &\xmark &\checkmark \\
\hline
Railgun~\cite{railgun} &\checkmark &\xmark &\xmark &\checkmark &\xmark &\xmark &\checkmark &\checkmark\\
\hline
Hinkal~\cite{hinkal} &\checkmark &\checkmark &\xmark &\xmark &\checkmark &\checkmark &\checkmark &\xmark \\
\hline
Privacy Pools~\cite{privacypools,buterin2023blockchain} &\checkmark &\checkmark &\xmark &\checkmark &\xmark &\checkmark &\checkmark &\checkmark \\
\hline
Nightfall~\cite{ey2025blockchain} &\xmark &\xmark &\xmark &\xmark &\xmark &\checkmark &\xmark &\xmark \\
\hline
\rowcolor{blue!10}
\textbf{Our framework} &\checkmark &\checkmark &\checkmark &\checkmark &\checkmark &\checkmark &\checkmark &\checkmark \\
\hline
\end{tabular}
\end{table}


% Symbols you can use to fill cells:
% \checkmark  (requires \usepackage{amssymb})
% \xmark (requires \usepackage{pifont}, then use \ding{55})
% $\circ$ for partial support
% Or just use: ✓, ✗, ◐



\section{Proposed Solution}

We propose an open framework where regulatory bodies publish compliance definitions, applications require users to prove they meet those requirements, and users generate cryptographic proofs demonstrating compliance, all without revealing private information.  Users who don’t meet the compliance requirements are unable to interact with compliant applications.  

Proving compliance to an application has been mentioned in brief in~\cite{} and is live in Railgun and Privacy Pools.  Our contribution is a standard framework to allow this compliance technique to be used at-scale across many applications and provides a standard interface for regulators to define compliance.

\subsection{Framework Approach}

The distinguishing feature of our solution relative to other work in this space is that it defines an open framework for use by players in a competitive ecosystem.  It is not a new blockchain.  It is not a payment system.  It is not a new application.  Rather, we are creating a framework for established and new-entrant financial providers to take advantage of the power of the emerging blockchain ecosystem by leveraging the features of our framework.

Our framework is unique in its provision of a standard means for regulators to publish requirements openly in a way that enables proof of compliance.  Regulation based on constraints (Section \ref{sec:constraints}) and proofs is valuable not only to financial providers, but also to regulators and users.  Regulators can check compliance by verification of proofs, which is much more efficient than having to evaluate full execution histories.  Users can show transparently that they are in compliance without having to reveal private data.

Our framework is \textit{chain agnostic}: it can be implemented on any chain with smart-contracts and capability to verify ZK proofs.  Because on-chain compliance pertains to transactions and transaction histories and not to specific proprietary features of a particular blockchain, our framework enables interoperability across platforms that use our framework.  This facilitates a competitive ecosystem in which each member of the ecosystem shares a common foundation of privacy and compliance guarantees.

\reminder{HFK: maybe a comment here about identity -- regulators have well-known
public keys and use private keys to sign regulations.  OR, reputable institutions
may translate regulations into constraints and sign those constraints.}

\subsection{Actors}
The actors in our framework consist of regulators, applications, and users.  A simplified overview of the actor interaction:

\begin{enumerate}
    \item \textit{Regulators} create and publish \textit{compliance definitions}.
    \item \textit{Applications} select relevant \textit{compliance definitions} based on their regulatory obligations.
    \item \textit{Users} generate ZK proofs of \textit{compliance definitions}, demonstrating that they meet requirements without revealing transaction histories, balances, or counterparties.
    \item \textit{Applications} verify proofs on-chain before allowing transactions, preventing non-compliant activity cryptographically rather than detecting it afterward.
\end{enumerate}

\subsubsection{Compliance Definition}

Compliance definitions in this framework are sets of logical rules over blockchain data with the goal of isolating non-compliant actors (e.g., ``address not on sanction list," ``funds from KYC'd exchange," ``no transaction-structuring patterns").  Compliance definitions contain public parameters, for example a list of sanctioned addresses, along with logical operations between those parameters.  Compliance definitions are more technically described in section \ref{sec:compliance-definition-in-constraint}.

A compliance definition consists of a ZK circuit and an on-chain proof verifier contract. 

Regulators (see below) create compliance definitions to enforce real-world regulation on-chain, for example a U.S. AML compliance definition.

Each published compliance definition can be used by any number of applications, and applications can require multiple compliance definitions.  Each government in the world has their own laws and regulations, and some applications wish to be compliant with multiple regulations.

Compliance definitions are updatable by their author to account for changing regulations.  For that reason, each compliance definition has an associated time parameter.

\subsubsection{Regulators}
\label{sec:regulators}
Regulators create compliance definitions over on-chain data and publish these definitions publicly with a signature.  The manner in which they are published is an implementation detail, but the compliance definitions need to be publicly accessible and in a form that can be compiled into a ZK circuit for users to prove.

Publishing a compliance definition is permissionless and we refer to any entity that does as a ``regulator".  Ideally, the entity is  associated with the government regulatory agency, but in practice, particularly initially, the publisher is most likely to be a blockchain analytics company.  These companies already assume compliance responsibilities, such as maintaining sanction lists and monitoring for suspicious chain activity.

As this is an open framework, there is no restriction on who can publish compliance definitions.  Applications need to be careful to select definitions published by reputable sources.

Regulators are also responsible for updating compliance definitions, like adding more constraints or updating parameters, with each update having a new
timestamp.  Here, ``update'' of a contract means a parameter change as
discussed in Section~\ref{sec:updateCompDef}.

\subsubsection{Applications}

On-chain applications determine which compliance definition(s) suit their regulatory obligations.

Applications integrate with this framework by making a smart-contract change to require users to provide a proof of the required compliance at entry-point functions.  If the proof does not verify, then the function reverts.  After this change, it is guaranteed that all future users of the application are compliant.

For example, the entrypoint function for some application is \texttt{deposit(someVars)}.  Integration would consist of changing it to add parameters \textit{proof} and the proof's required inputs \textit{publicInputs}: \texttt{deposit(someVars, proof, publicInputs)}, and then beginning the function by using a \textbf{require} statement to mandate verification of the proof: \texttt{require(verifierContract.verify(proof, publicInputs))}.

\subsubsection{Users}

Users locally generate and store ZK proofs of compliance definitions and supply the proofs as an argument when interacting with a compliance requiring application.  The constraint system was constructed to intelligently minimize the amount of proof computation needed for users interacting with many compliance requiring applications.  Users are resposible for locally keeping previously generated proofs.  This process is aided by the proof manager, discussed in section \ref{sec:proof-manager}.

\reminder{HFK: Users have to be able to state what sort of compliance they claim
to prove.  Thus, constraints need to have unique publicly-available IDs.  
the result is that we have a set of identified constraints, each signed by
an identified regulator.  This database needs to be maintained in a trustless
manner and be accessible.  Control of this database is another censorship risk.}
\reminder{Joss: We could deploy a ``register" contract on-chain that regulators must get an id when they deploy or update a compliance definition, but this would create a trust dependency because some entity would have to deploy this ``register" on each new chain.  In the event 2 ``registers" are deployed by independent actors, how do applications and regulators resolve which is the canonical register?  I think we can avoid this dependency all together.  There can be a field in the verifier contract that is an IPFS hash of the circuit for this verifier.  Then users can always look up the circuit given the verifier contract address.}

\reminder{HFK: we need a discussion here -- may a new subsubsection -- on user identity.  Our framework permits users to be fully anonymous while proving compilance, but the particular application may not be set up for total anonymity.  If the application keeps records based on users, there will be some sort of identity in the application, either the user's real identity or a pseudonym.  There is, of course, some loss of privacy here.  Such a loss may be unacceptable in some settings, but in other cases, it may be part of the required business practices. 

Consider a stablecoin provider. Transactions could be totally public.  A better, though not perfect solution is to have deposits and payments of fungible stablecoins with balances held internally in the stablecoin provider's database/contract.  The provider could keep only balances, not transaction
histories.  This would enable some privacy loss and some forms of censorship, depending on how anonymous the user is.  Alternatively, the user could hold the balance and prove as needed that the balance is large enough.    While the latter may be better from a privacy standpoint, the former is closer to many current system structures and might be a reasonable first step as stablecoins compete on privacy.

I'm running off at the mouth here.  Edit as you see fit!!!

Another point: compliance constraints are separate from the user and the application.  Proofs can relate not only to compliance constraints but to application-specific features.

}
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=1.5cm and 2cm,
    user/.style={rectangle, draw=gray!70, fill=gray!20, thick, minimum width=1.5cm, minimum height=1.2cm, rounded corners=5pt, align=center, font=\normalsize},
    compliance/.style={rectangle, draw=red!70, fill=red!10, very thick, minimum width=3cm, minimum height=1.5cm, rounded corners=10pt, align=center, font=\normalsize},
    dapp/.style={rectangle, draw=cyan!70, fill=cyan!10, very thick, minimum width=3cm, minimum height=1.5cm, rounded corners=10pt, align=center, font=\normalsize},
    arrow/.style={->, >=stealth, thick, black},
    dashedarrow/.style={->, >=stealth, thick, dashed, cyan!70},
    regarrow/.style={->, >=stealth, thick, green!60!black},
    label/.style={font=\scriptsize, text=black}
]

% User node (left)
\node[font=\normalsize] (user) {User};

% Compliance boxes
\node[compliance, right=3cm of user, yshift=1cm] (uscomp) {Compliance\\definition A};
\node[compliance, right=3.5cm of uscomp] (eucomp) {Compliance\\definition B};

% Regulator arrow origins (invisible nodes above compliance boxes)
\coordinate[above=1.5cm of uscomp] (reg1);
\coordinate[above=1.5cm of eucomp] (reg2);

% Regulator labels at the top
\node[above=0.1cm of reg1, font=\normalsize, text=green!60!black] {Regulator};
\node[above=0.1cm of reg2, font=\normalsize, text=green!60!black] {Regulator};

% Dapp boxes (bottom)
\node[dapp, below=2.5cm of uscomp, xshift=-0.5cm] (dapp1) {Dapp 1};
\node[dapp, below=2.5cm of eucomp, xshift=0.5cm] (dapp2) {Dapp 2};

% Arrows from regulators to compliance boxes
\draw[regarrow] (reg1) -- (uscomp) node[midway, right, label, text=green!60!black] {publishes};
\draw[regarrow] (reg2) -- (eucomp) node[midway, right, label, text=green!60!black] {publishes};

% Arrow from user to US compliance
\draw[arrow] (user) -- (uscomp) node[midway, above=0.1cm, label, align=center] {generates\\proof of};

% Arrow from user to dapps area
\draw[arrow] (user) -- (dapp1.north west) node[midway, left=0.1cm, label] {proof validates};

% Dashed arrows from compliance to dapps
\draw[dashedarrow] (dapp1) -- (uscomp) node[midway, left, label, text=cyan!70] {requires};
\draw[dashedarrow] (dapp2) -- (uscomp) node[midway, left=0.15cm, label, text=cyan!70] {requires};
\draw[dashedarrow] (dapp2) -- (eucomp) node[midway, right, label, text=cyan!70] {requires};

\end{tikzpicture}
\caption{Composable compliance framework. Applications can opt into requiring multiple compliance definitions.  Users can generate proofs of any compliance definition.  In this example, a user is interacting with Dapp 1 so they must first generate a proof of compliance definition A.}
\label{fig:compliance-framework-overview}
\end{figure}


\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=2cm and 2.5cm,
    user/.style={circle, draw=white, fill=black!80, thick, minimum size=1.3cm, align=center, font=\normalsize, text=white},
    compliance/.style={rectangle, draw=red!70, fill=red!10, very thick, minimum width=3cm, minimum height=1.5cm, rounded corners=10pt, align=center, font=\normalsize},
    dapp/.style={rectangle, draw=cyan!70, fill=cyan!10, very thick, minimum width=3cm, minimum height=1.5cm, rounded corners=10pt, align=center, font=\normalsize},
    arrow/.style={->, >=stealth, thick, black},
    dashedarrow/.style={->, >=stealth, thick, dashed, cyan!70},
    label/.style={font=\scriptsize, text=black}
]

% User node (top center)
\node[font=\normalsize] (user) {User};

% Compliance boxes (left and right of user, slightly below)
\node[compliance, below left=1.5cm and 3cm of user] (uscomp) {Compliance\\definition A};
\node[compliance, below right=1.5cm and 3cm of user] (eucomp) {Compliance\\definition B};

% Dapp box (bottom center)
\node[dapp, below=4cm of user] (dapp) {Dapp};

% Arrows from compliance boxes to user
\draw[arrow] (user) -- (uscomp) node[midway, above, label, sloped, align=center] {generates\\proof of};
\draw[arrow] (user) -- (eucomp) node[midway, above, label, sloped, align=center] {generates\\proof of};

% Arrow from user to dapp
\draw[arrow] (user) -- (dapp) node[midway, right, label, align=center] {both\\proofs validate};

% Dashed arrows from compliance boxes to dapp
\draw[dashedarrow] (dapp) -- (uscomp) node[midway, left=0.15cm, label, text=cyan!70] {requires};
\draw[dashedarrow] (dapp) -- (eucomp) node[midway, right=0.15cm, label, text=cyan!70] {requires};

\end{tikzpicture}
\caption{Example compliance scenario. A Dapp requires both A and B compliance. The user generates proofs for both compliance definitions and submits them together. The Dapp verifies both proofs on-chain before allowing the user to interact.}
\label{fig:multi-compliance-example}
\end{figure}

\subsection{Composability} 
Because compliance definitions are public, the same definition can be used by any number of protocols, a user can generate proofs for any published definitions of compliance, and proofs for multiple compliance definitions can be required by a protocol.  Figure~\ref{fig:compliance-framework-overview} shows dapps opting into any number of compliance definitions.  Figure ~\ref{fig:multi-compliance-example} shows a user proving compliance for a dapp that requires multiple compliance definitions.

Compliance definitions can be fine-grained instead of wide-umbrella definitions.  For example, instead of having a single definition that encompasses all US rules and regulations, there can be a definition for US consumer compliance and a definition for US corporate compliance.

The cost to verify a proof is independent of what is being proved.  This is inherited from using zk-SNARKs as compliance proofs.  The circuit can be very computationally intensive and have pages of criteria in its natural language form, but, although proof generation is computationally intensive, verification is not.  Ease of verification enables efficient composability.

\subsection{Constraints}
\label{sec:constraints}
Although regulations vary globally and over time, they often rely on similar logical primitives: checking membership in lists, comparing values against thresholds, or verifying historical patterns. We built our framework to leverage this similarity by decomposing compliance definitions into reusable logic that we refer to as \textit{constraints}.  This compositional structure allows for complex requirements to be expressed as combinations of constraints.
 
A \textit{constraint} is a boolean predicate that evaluates blockchain state using first-order logic. A constraint $C$ is expressed in the general form:

$$C = Q \, x \in D(s, i) : P(x, p)$$

where:
\begin{itemize}
    \item \textbf{Quantifier} ($Q \in \{\forall, \exists, \ldots\}$): the logical quantifier

    \item \textbf{Chain State} ($i$): the blockchain state(s) from which the domain is evaluated.  Chain states can be:
    \begin{itemize}
        \item Atomic states:
        \begin{itemize}
            \item Single block: $B_k$
            \item Block range: $[B_i, B_j]$
        \end{itemize}
        \item State sets (with implicit universal quantification):
        \begin{itemize}
            \item Parameterized windows: $\{[B_i, B_j] \subseteq [B_0, B_{\text{current}}] : \text{condition}\}$
        \end{itemize}
    \end{itemize}
    
    \item \textbf{Domain} ($D(s, i) \rightarrow \{x, \ldots\}$): a function that maps a subject $s$ and chain state $i$ to a collection of blockchain entities $x$ over which the quantifier ranges. Common domains include:
    \begin{itemize}
        \item Singleton: $D(s, i) = \{s\}$ evaluates a single subject (atomic case)
        \item Transaction history: $D(\text{addr}, i) = \text{history}(\text{addr}, i)$ 
        \item Set of recipients: $D(\text{addr}, i) = \text{recipients}(\text{addr}, i)$
        \item Filtered collections: $D(s, i) = \{x \in D'(s, i) : \text{filter}(x)\}$
    \end{itemize}
    
    \item \textbf{Subject} ($s$): the root blockchain entity from which the domain is derived. Example subjects include:
    \begin{itemize}
        \item Sender of a transaction: $\text{sender}(\text{txn})$
        \item Recipient of a transaction: $\text{recipient}(\text{txn})$
        \item Account balance: $\text{balance}(\text{addr})$
        \item Result of a contract call: $\text{result}(\text{contract fn})$
        \item Transaction payload: $\text{payload}(\text{txn})$
    \end{itemize}

    \item \textbf{Parameter} ($p$): the value or set against which elements are evaluated, containing:
    \begin{itemize}
        \item Public parameters $p_{\text{pub}}$ provided by the compliance author (e.g., sanction lists, required values)
        \item Private inputs $p_{\text{priv}}$ provided by the user (e.g., mixer secret, identifying information)
    \end{itemize}
    
    \item \textbf{Predicate ($P \rightarrow \{\text{true}, \text{false}\}$)}: a boolean function applied to domain elements. Predicates can operate on:
    \begin{itemize}
        \item Individual elements ($P(x, p)$, where $x$ is an element in the set returned by $D(s, i)$):
        \begin{itemize}
            \item Comparison operations: $P(x, p) = (x \,\phi\, p)$ where $\phi \in \{=, \neq, <, \leq, >, \geq, \in, \notin, \ldots\}$
            \item Nested constraints over the same chain state: $P(x, p) = C'(x, p', i')$
        \end{itemize}
        \item Entire collections ($P(X, p)$), where $X$ is a subset of the set returned by $D(s, i)$:
        \begin{itemize}
            \item Aggregate comparisons: $P(X, p) = (\text{agg}(X) \,\phi\, p)$ where $\text{agg} \in \{\sum, \text{count}, \text{max}, \text{min}, \ldots\}$
        \end{itemize}
    \end{itemize}
\end{itemize}

\subsubsection{Predicate constraints as special cases}
Often regulators will want to make constraints independent of historic chain-state.  These are useful for simple statements like ``This address is not on sanction list $A$", where $A$ is a public parameter available at the current chain state.

To simplify notation, we introduce a subclass of constraints called \textit{predicate constraints}.  Predicate constraints function exactly the same as the above constraint but with simplified syntax:

$$C = s \,\phi\, p$$

where $\phi \in \{=, \neq, <, \leq, >, \geq, \in, \notin, \ldots\}$, $p$ is the parameter passed in, and $s$ is the subject being evaluated.  When predicate constraint notation is used, the chain state can be assumed to be the current block.  

\subsubsection{Example constraints}

\paragraph{Simple example:}
Sanction list checks are a popular compliance measure in the blockchain space~\cite{solomka2025zeroknowledge, tornadocashprimer}.  The constraint ``Sender of this transaction is not on sanction list $A$" can be decomposed as the following predicate constraint:

\begin{itemize}
    \item \textbf{Subject}: $\text{sender}(\text{txn})$ --- the address that initiated the transaction
    \item \textbf{Operation}: $\notin$ --- does not exist in
    \item \textbf{Parameter}: $A$ --- the sanction list (a public parameter provided by the regulator)
\end{itemize}

The constraint evaluates to \textbf{true} if and only if the subject (sender of the transaction) does not exist in the parameter (sanction list).  Formally expressed as:

$$C = \text{sender}(\text{txn}) \notin A$$

\paragraph{Full example:} 
The constraint ``Throughout their transaction history, the sender of this transaction never interacted with a specific protocol" demonstrates the full constraint form.  It can be decomposed as follows:

\begin{itemize}
    \item \textbf{Quantifier}: $\forall$ --- universal quantifier; the predicate must hold for all elements in the domain
    \item \textbf{Chain State}: $[B_0, B_{\text{current}}]$ --- from genesis block to current block
    \item \textbf{Subject}: $\text{sender}(\text{txn})$ --- the address that initiated the transaction
    \item \textbf{Domain}: $\text{txn\_history}(\text{sender}(\text{txn}), [B_0, B_{\text{current}}])$ --- all transactions sent by the subject within the chain state
    \item \textbf{Parameter}: $p_{\text{protocol}}$ --- the address of the prohibited protocol (a public parameter provided by the regulator)
    \item \textbf{Predicate}: $\text{recipient}(\text{t}) \neq p_{\text{protocol}}$ --- the recipient of each historical transaction is not the prohibited protocol
\end{itemize}

The constraint evaluates to \textbf{true} if and only if every transaction in the sender's history was sent to an address other than the prohibited protocol.  Formally expressed as:

$$C = \forall \, \text{t} \in \text{txn\_history}(\text{sender}(\text{txn}), [B_0, B_{\text{current}}]) : \text{recipient}(\text{t}) \neq p_{\text{protocol}}$$

\bigskip
For more examples, see section \ref{sec:prototype-constraints} where we define the constraints used in our prototype implementation.

\subsubsection{Compliance definition in constraint language} 
\label{sec:compliance-definition-in-constraint}
A compliance definition is a combination of constraints with boolean operators and a time parameter during which the definition is valid.   Given constraints $C_1, \ldots, C_k$, a compliance definition $R$ is defined as:
$$R = C_1 \circ C_2 \circ \cdots \circ C_k, [t_{\text{start}}, t_{\text{end}}]$$
where $\circ$ denotes arbitrary boolean operators ($\land, \lor, \lnot$, and parentheses for grouping), allowing for complex logical combinations of constraints. The time parameters $t_{\text{start}}$ and $t_{\text{end}}$ denote the block heights or timestamps defining the validity period of the compliance definition. This temporal component is necessary to account for regulations changing over time and to allow for regulations to be published before they go into effect.

The definition $R$ is considered active if and only if the evaluation occurs within this temporal window:
$$\text{active}(R, t) = \begin{cases}
\text{true} & \text{if } t_{\text{start}} \leq t \leq t_{\text{end}} \\
\text{false} & \text{otherwise}
\end{cases}$$

Alternatively, open-ended validity can be expressed with $t_{\text{end}} = \infty$ for rules with no expiration.  For brevity, the time parameter is often omitted in this paper for compliance definitions and can be assumed to be valid for the current time.

We envision compliance definitions being able to be compiled into constraints, and constraints being able to compile into circuits.  This allows compliance to be proven by proving the constraints that make up a definition.

\subsubsection{Example: shared constraints}
\label{sec:shared-constraints}
When multiple compliance definitions require the same constraint, that constraint can be proved once and applied to multiple definitions.  For example, an application requires compliance definitions $R_1$ and $R_2$ which are made up of constraints $C_1$ and $C_2$:

$$R_1 = C_1 \land C_2$$
$$R_2 = C_1$$

A user of the application only needs to compute a proof for $R_1$ because its constraints are a super-set of the constraints used in $R_2$: proving $R_1$ also proves $R_2$.

\paragraph{Another Example}
An application requires $R_1$ and $R_2$:
$$R_1 = C_1 \land C_2$$
$$R_2 = C_2 \land C_3$$

Constructing a proof that validates for both regulations only requires generating a proof for three constraints, as $C_2$ is included in both definitions.  $C_2$ is proved once and that proof is applied to both $R_1$ and $R_2$.


\subsubsection{Off-chain data in constraints}
Constraints can require off-chain data that is provided by the user during proof generation.  This user input, $p_{\text{priv}}$, is completely private and not revealed to any other entity.  This allows for constructing constraints that rely on off-chain data.

Public off-chain data, such as sanction lists, can also be used as constraints by writing the data on-chain.

\subsection{Compliance definition contract}

The \texttt{ComplianceDefinition} contract serves as the stable on-chain representation of a compliance definition. Applications reference this contract's address, which remains constant even as the underlying verification logic is updated.

Each update to a compliance definition, whether a constraint update or a public parameter update, is recorded as a new \texttt{ComplianceVersion}. This provides a complete audit trail and enables verification against any historical state of the definition.

\begin{codebox}
\Procname{Contract: ComplianceDefinition}
\li \kw{struct} $\id{ComplianceVersion}$
\li \>\>$\id{verifier}$ (address of verifier contract for this version)
\li \>\>$\id{paramsRoot}$ (On-chain commitment of public parameters)
\li \>\>$\id{tStart}$ (block height when this version becomes active)
\li \>\>$\id{tEnd}$ (block height when this version expires
\li \>\>$\id{metadataHash}$ (ex: IPFS hash of compliance circuit and full public parameters)
\li
\li \Comment State variables
\li $\id{versions}$ (array of $\id{ComplianceVersion}$)
\li $\id{regulator}$ (authorized address for updates)
\li
\li $\proc{Verify}(\id{proof, publicInputs})$
\li \>$\id{v} \gets \proc{GetActiveVersion}()$
\li \>\Return $\id{v.verifier}.\proc{VerifyProof}(\id{proof}, \id{publicInputs}, \id{v.paramsRoot})$
\li
\li $\proc{UpdateConstraint}(\id{newVerifier}, \id{newParamsRoot}, \id{tStart}, \id{tEnd}, \id{metadataHash})$
\li \>\kw{require} $\id{msg.sender} = \id{regulator}$
\li \>$\id{versions}.\proc{push}(\id{ComplianceVersion}\{$
\li \>\>\>$\id{newVerifier}, \id{newParamsRoot}, \id{tStart}, \id{tEnd}, \id{metadataHash}\})$
\li
\li $\proc{UpdateParams}(\id{newParamsRoot}, \id{tStart}, \id{tEnd}, \id{metadataHash})$
\li \>\kw{require} $\id{msg.sender} = \id{regulator}$
\li \>$\id{currentVerifier} \gets \proc{GetActiveVersion}().\id{verifier}$
\li \>$\id{versions}.\proc{push}(\id{ComplianceVersion}\{$
\li \>\>\>$\id{currentVerifier}, \id{newParamsRoot}, \id{tStart}, \id{tEnd}, \id{metadataHash}\})$
\li
\li $\proc{GetActiveVersion}() \rightarrow \id{ComplianceVersion}$
\li \>\Return most recent version where $\id{tStart} \leq \id{block.number} \leq \id{tEnd}$
\li
\li $\proc{GetVersionAt}(\id{blockHeight}) \rightarrow \id{ComplianceVersion}$
\li \>\Return version where $\id{tStart} \leq \id{blockHeight} \leq \id{tEnd}$
\li
\li $\proc{GetVersionCount}() \rightarrow \id{uint}$
\li \>\Return $\id{versions}.\id{length}$
\end{codebox}

The \texttt{ComplianceVersion} struct captures the complete state of a compliance definition at any point in time:

\begin{itemize}
    \item \texttt{verifier} The ZK verifier contract containing the verification key for the circuit. Changes only on constraint updates.
    \item \texttt{paramsRoot} Merkle root commitment to public parameters (e.g., sanction list). The full parameter set is stored off-chain and referenced via \texttt{metadataHash}.
    \item \texttt{tStart}, \texttt{tEnd} Block heights defining the validity window, enabling compliance definitions to be published before they take effect.
    \item \texttt{metadataHash} IPFS content hash linking to the constraint DSL specification, full public parameters, and regulator signature.
\end{itemize}

This design provides several benefits: (1) a complete audit trail of all compliance updates, (2) the ability to verify proofs against historical definitions, (3) support for publishing updates before they become active, and (4) efficient on-chain storage by committing only the Merkle root of potentially large parameter sets.

The verifier contract is an upgradable contract deployed and updated by its regulator and keeps a history of all compliance definitions it has previously verified for by referencing the addresses and IDs of each definition's corresponding verifier contract.  The individual verifiers contain the time bounds $t_start$ and $t_end$ that annotate when that compliance definition is considered active.   

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=1.5cm,
    constraint/.style={rectangle, draw=blue!60, fill=blue!5, thick, minimum width=3cm, minimum height=1cm, rounded corners},
    proof/.style={rectangle, draw=green!60, fill=green!5, thick, minimum width=2.5cm, minimum height=0.8cm, rounded corners},
    aggregate/.style={rectangle, draw=purple!60, fill=purple!5, thick, minimum width=4cm, minimum height=1.2cm, rounded corners},
    onchain/.style={rectangle, draw=orange!60, fill=orange!5, thick, minimum width=4cm, minimum height=1cm, rounded corners},
    arrow/.style={->, >=stealth, thick},
    label/.style={font=\small}
]

% Layer 1: Individual Constraints
\node[constraint, align=center] (c1) {$C_1$};
\node[constraint, align=center, right=of c1] (c2) {$C_2$};
\node[constraint, align=center, right=of c2] (c3) {$C_3$};

% Layer 2: Individual Proofs
\node[proof, below=of c1] (p1) {Proof $\pi_{C_1}$};
\node[proof, below=of c2] (p2) {Proof $\pi_{C_2}$};
\node[proof, below=of c3] (p3) {Proof $\pi_{C_3}$};

% Arrows from constraints to proofs
\draw[arrow] (c1) -- node[right, label] {prove} (p1);
\draw[arrow] (c2) -- node[right, label] {prove} (p2);
\draw[arrow] (c3) -- node[right, label] {prove} (p3);

% Layer 3: Aggregation (increased spacing)
\node[aggregate, align=center, below=3.5cm of c2] (agg) {Aggregated Proof\\$\pi_R = \text{Aggregate}(\pi_{C_1}, \pi_{C_2}, \pi_{C_3})$};

% Arrows from proofs to aggregation
\draw[arrow] (p1) -- (agg.north west);
\draw[arrow] (p2) -- (agg.north);
\draw[arrow] (p3) -- (agg.north east);

% Layer 4: On-chain Verification
\node[onchain, align=center, below=of agg] (verify) {On-Chain Verification\\{\scriptsize \texttt{verifierContract.verify}($\pi_R$)}};

% Arrow from aggregation to on-chain
\draw[arrow, very thick] (agg) -- node[right, label] {submit} (verify);

% Compliance Definition Label
\node[above=0.3cm of c2, font=\large\bfseries] (title) {Compliance Definition: $R = C_1 \land C_2 \land C_3$};

% Layer Labels
\node[left=1.5cm of c1, font=\small\itshape, align=right] (layer1) {Constraints};
\node[left=1.5cm of p1, font=\small\itshape, align=right] (layer2) {Constraint Proofs};
\node[left=1.5cm of agg, font=\small\itshape, align=center] (layer3) {Aggregated Proof\\(Compliance Definition Proof)};
\node[left=1.5cm of verify, font=\small\itshape, align=right] (layer4) {Verification};

% Background boxes for visual grouping
\begin{scope}[on background layer]
    \node[draw=blue!30, fill=blue!3, thick, rounded corners, fit=(c1)(c2)(c3), inner sep=0.3cm] {};
    \node[draw=green!30, fill=green!3, thick, rounded corners, fit=(p1)(p2)(p3), inner sep=0.3cm] {};
    \node[draw=purple!30, fill=purple!3, thick, rounded corners, fit=(agg), inner sep=0.4cm] {};
\end{scope}

\end{tikzpicture}
\caption{Proof aggregation workflow for compliance definition $R$. Individual constraints are proved separately, generating proofs $\pi_{C_1}$, $\pi_{C_2}$, and $\pi_{C_3}$. These proofs are then aggregated into a single proof $\pi_R$ that is submitted on-chain for verification.}
\label{fig:proof-aggregation-layered}
\end{figure}

\subsection{Generating Compliance Proofs}

Proofs of compliance definition are aggregate proofs of multiple constraint proofs.  When proving a compliance definition, each constraint is proved individually, then all constraint proofs are combined in an aggregation proof which is then submitted on-chain.  See Figure~\ref{fig:proof-aggregation-layered} for the complete workflow.

Proofs are generated locally by the user and then submitted on-chain as transaction arguments.  The proof is public on-chain\footnote{Some blockchains allow for private transactions.  Proofs submitted as part of a private transaction will not be public.}, but it cannot be used to recover the user's private inputs.

The user can locally save all constraints that they have proven, discussed further in section \ref{sec:proof-manager}.  This allows users to re-use previous proved constraints.  It is often the case that some parts of a compliance definition may already be proved or that the constraints of separate regulatory domains have some subexpressions in common.  In these cases, inference over the constraints can lead to significant reduction in the complexity of the constraints to be proved, shown in section \ref{sec:shared-constraints}.

\paragraph{Example} 
Alice previously proves compliance definition $R_1$, which requires proving constraints $C_1$ and $C_2$.  Alice has locally saved her proofs of $C_1$ and $C_2$.

$$R_1 = C_1 \land C_2$$

Now, Alice wishes to prove $R_2$.  
 
$$R_2 = C_2 \land C_3$$

She can inspect $R_2$ and sees that it uses $C_2$, a constraint that she has previously proved and saved.  She can skip the computation of $C_2$ because she already has a valid proof!  All she needs to compute is $C_3$, and $R_2$, which is the aggregate proof over $C_2$ and $C_3$.

\subsection{Proof Manager}
\label{sec:proof-manager}
 We envision a client-side program that aids in proof generation as seen in figure \ref{fig:proof-manager-architecture}. The proof manager takes as input a compliance definition, chain identifier, private user input, and the user address.  It outputs a proof of that compliance definition.

The proof manager is responsible for
\begin{itemize}
    \item Doing inference over required constraints to identify the minimum amount of proving necessary, relying on previously proved constraints.  Minimizing proofs needed is more efficient than speeding up proof generation.
    \item Constraint to circuit compilation.
    \item Fetching on-chain data of the user's transaction history necessary for proof generation.
    \item Fetching public proof inputs from the published verifier contract.
    \item Generating constraint proofs.
    \item Generating aggregate proof over the constraints to prove compliance.
    \item Saving completed proofs.
\end{itemize}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=0.8cm and 1.5cm,
    input/.style={rectangle, draw=blue!70, fill=blue!15, thick, minimum width=2.3cm, minimum height=0.65cm, rounded corners, align=center, font=\small},
    process/.style={rectangle, draw=green!70, fill=green!20, thick, minimum width=2.3cm, minimum height=0.7cm, rounded corners, align=center, font=\small},
    storage/.style={rectangle, draw=brown!70, fill=brown!20, thick, minimum width=2cm, minimum height=1cm, rounded corners, align=center, font=\small},
    output/.style={rectangle, draw=orange!70, fill=orange!20, thick, minimum width=3.2cm, minimum height=0.85cm, rounded corners, align=center, font=\normalsize},
    manager/.style={rectangle, draw=red!60, fill=red!8, very thick, rounded corners},
    blockchain/.style={rectangle, draw=cyan!70, fill=cyan!8, very thick, rounded corners},
    onchainbox/.style={rectangle, draw=cyan!60, fill=cyan!15, thick, minimum width=2.8cm, minimum height=0.9cm, rounded corners, align=center, font=\small},
    arrow/.style={->, >=stealth, thick},
    dashedarrow/.style={->, >=stealth, thick, dashed},
    label/.style={font=\tiny, text=black}
]

% Input nodes (left side)
\node[input] (in1) {Compliance\\definition $R$};
\node[input, below=0.25cm of in1] (in2) {Chain identifier};
\node[input, below=0.25cm of in2] (in3) {Verifier contract\\address};
\node[input, below=0.25cm of in3] (in4) {User address};
\node[input, below=0.25cm of in4] (in5) {Private input};

% Proof Manager container
\node[manager, right=1.5cm of in3, minimum width=7.3cm, minimum height=8cm] (manager) {};
\node[above=0.05cm of manager.north, font=\bfseries\large, text=red!70] {Proof Manager};

% Internal components - circuit moved 0.1cm to the right (xshift changed from 1.5cm to 1.6cm)
\node[process, below=0.7cm of manager.north, xshift=-1.8cm] (constraint) {Constraint\\inference};
\node[process, below=0.7cm of manager.north, xshift=1.7cm] (circuit) {Circuit\\compilation};
\node[storage, below=1.9cm of constraint] (storage) {Proof\\storage};
\node[process, below=1.2cm of circuit] (fetchdata) {Fetch on-chain\\data};
\node[process, below=1.2cm of fetchdata] (generate) {Generate\\constraint\\proofs};
\node[process, below=2.2cm of storage] (aggregate) {Aggregate\\proofs};

% Blockchain container (right side)
\node[blockchain, right=2.2cm of fetchdata, minimum width=3.5cm, minimum height=4.5cm] (blockchain) {};
\node[above=0.05cm of blockchain.north, font=\bfseries\large, text=cyan!70] {Blockchain};

\node[onchainbox, below=0.6cm of blockchain.north] (txhistory) {User transaction\\history};
\node[onchainbox, below=1.6cm of txhistory] (verifier) {Verifier contract\\(public inputs)};

% Output (centered below aggregate proofs)
\node[output, below=1.5cm of aggregate] (output) {Aggregated proof $\pi_R$};

% Input arrows to manager
\draw[arrow] (in1.east) -- (manager.west |- in1.east);
\draw[arrow] (in2.east) -- (manager.west |- in2.east);
\draw[arrow] (in3.east) -- (manager.west |- in3.east);
\draw[arrow] (in4.east) -- (manager.west |- in4.east);
\draw[arrow] (in5.east) -- (manager.west |- in5.east);

% Internal workflow arrows
\draw[arrow] (constraint) -- (storage) node[midway, right, label, align=center] {skip generation if\\previously proved};
\draw[arrow] (circuit) -- (fetchdata);
\draw[arrow] (fetchdata) -- (generate);

% Save proof arrow - STRAIGHT LINE from generate to storage
\draw[arrow] (generate.west) -- (storage.south east) node[pos=0.8, right=0.05cm, label] {save proof};

% Load proof from storage to aggregate
\draw[arrow] (storage) -- (aggregate) node[midway, right, label, align=left] {load\\proofs};

% Blockchain interaction arrows
\draw[dashedarrow, cyan!70] (fetchdata.east) -- (txhistory.west);
% STRAIGHT LINE to verifier contract
\draw[dashedarrow, cyan!70] (fetchdata.east) -- (verifier.west);

% Output arrow
\draw[arrow, very thick] (aggregate.south) -- (output.north) node[midway, right, label] {Output};

% Input label
\node[left=0.3cm of in3, font=\small\bfseries, text=blue!70] {Inputs};

\end{tikzpicture}
\caption{Proof Manager architecture. The system processes inputs through constraint inference (which checks proof storage for previously generated proofs), circuit compilation, fetching on-chain data from both the blockchain transaction history and verifier contract, and proof generation. Generated proofs are saved to proof storage and then loaded by the aggregation step to produce the final compliance proof $\pi_R$.}
\label{fig:proof-manager-architecture}
\end{figure}

\subsection{Updating Compliance Definition}
\label{sec:updateCompDef}
Laws and regulations are constantly updated by governments; therefore, definitions of compliance will change over time.  The framework allows updates to compliance definitions.

A compliance definition can be updated in two ways, and both cases involve the regulator changing the \texttt{ComplianceDefinition} contract state.
\begin{itemize}
    \item Public parameter $p_{\text{pub}}$ update (ex: add or remove an address from the sanction list) modifies the public inputs used in proof verification
    \item Constraint update (ex: add or remove a constraint from the definition) modifies the circuit.  A constraint update also requires regulators to publicly re-publish and sign the new compliance definition circuit.
\end{itemize}

Although our framework is independent of the specific prover used, there is a clear need to use a prover with either no trusted setup or a reusable trusted setup, for example PLONK~\cite{Gabizon2019PLONKPO}. The number of distinct proofs that need to be computed in any transaction system using our framework makes it highly impractical to perform circuit-specific setup ceremonies. The options for handling updates are to deploy a new contract or update the existing contract.

\begin{itemize}
    \item Regulator publishes a new verifier contract for the updated compliance and \textbf{all} applications must update their smart-contracts to reference this new contract address in order to remain compliant.
    \item Regulator updates their current verifier contract and no change is needed by the application.
\end{itemize}

We investigate both these approaches and find that verifier contract updating is better than deploying a new verifier contract.

 If regulators publish a new verifier contract each time a sanction list is updated for a compliance definition, the applications using that compliance definition must manually update their smart contracts to reflect the new verifier contract address.  As an example, Chainalysis, a leading blockchain analytics company used by government bodies, adds addresses to their on-chain sanctions list roughly every 30 days~\cite{chainalysissanctionsoracle}.  It is infeasible to require all applications to update their smart contract every 30 days.  A priority of this framework is to provide applications with a strong guarantee of compliance and minimal overhead to the application.

For these reasons, we choose the path of regulators updating their existing compliance contract instead of publishing a new one on each update.  This will keep the compliance verifier contract at the same address and does not require any action from the application.  With this technique, applications have the guarantee that all their users meet the compliance definition with no windows of non-compliance.

It should be noted that updating sanction lists can introduce a brief period of regulatory arbitrage, as discussed in~\cite{hazedaze,buterin2023blockchain}.  This is the period after an address does something illegal, like participating in an exploit or money laundering, but before sanction lists have been updated with their addresses. Currently, blockchain analytics companies are responsible for regularly updating sanctions lists.  These blockchain analytics companies compete on speed and accuracy of their sanction list updates.  This framework allows compliance definitions' sanction lists to be updated by these analytics companies, which aligns with their current responsibilities and strengthens the incentive for fast updates.  Applications that want even stronger guarantees can also implement a deposit delay that requires users to prove compliance for a period after deposit.

\subsubsection{Post Update Re-proving}

Users who generated but did not use a compliance proof before the regulator updated the verifier contract may have to re-generate parts of their proof.  They will have to regenerate the aggregate proof over the constraints, but might not have to re-prove all constraints in that compliance definition.  If a definition consists of constraints $C_1$ and $C_2$, and the regulator updates the public parameters of $C_1$, users will have to reprove $C_1$ but are able to re-use their proof of $C_2$.

Re-proving only the constraints that changed in an update is less costly than re-proving the entire compliance definition.  Regulators have some incentive to make small updates since ease of proof generation leads to ease of compliance.

\subsubsection{Cross-chain Update Considerations}
If a compliance definition is updated across multiple independent blockchains, there can exist a small window of time after the update is finalized on one chain but before it's finalized on another.  There is no native solution for atomic cross-chain transactions that accounts for different data layers and virtual machines. 

The solution is to publish the updated compliance before it goes into effect.  Recall that the compliance definition has time bounds: $$R = C_1 \land C_2 \land \cdots \land C_k, [t_{\text{start}}, t_{\text{end}}]$$
Compliance definitions can be published before $t_\text{start}$ time.  If it is published far enough ahead of $t_\text{start}$, it is likely that the update transaction will be finalized on all chains by $t_\text{start}$ when the update goes into effect.  This is a natural consequence of making time a first-class parameter in compliance definitions.

\subsection{Regulator Trust Assumption}
Because compliance is updatable, there exists a risk vector of the regulator of a compliance definition turning malicious and publishing an unjust or incorrect update.  For example, a regulator adds an address to the sanction list to censor a particular law-abiding citizen.

This framework does not prevent against malicious regulators, but it provides a solution.  A key property of this framework is that publishing compliance definitions is \textbf{permissionless}.  If an application is not satisfied with any of the available compliance definitions or untrusting of the regulators, they can easily publish and maintain their own definition.  They can create a compliance definition from scratch or fork one of the publicly available definitions.  However, ``in-house" compliance results in more work for the application's team.

Additionally, because compliance definitions are public, any malicious update is also public.  Transparent regulation is much safer for users and applications because the regulator must weigh the benefit of the malicious action against the consequences of the public fallout.

\section{Prototype}
This framework is applicable to any blockchain with a smart-contract language and constraints can be written in any syntax as long as they are representable in a ZK circuit.  For our initial prototype and benchmarking of this framework, constraints are expressed in the Noir ZK DSL~\cite{AztecNoir22} with the Barretenberg backend to generate PLONKish\footnote{Specifically, we generate UltraHonk proofs, which is a variation of PLONK.} proofs.  The application integrating this framework will use the Ethereum Virtual Machine (EVM) expressed through the Solidity smart-contract DSL.  Noir was chosen because it provides high-level language semantics for writing ZK circuits, but any ZK circuit language can be used.  The EVM and Solidity were chosen because they have a large and polished ecosystem of smart-contract tools.

\subsection{Prototype Constraints}
\label{sec:prototype-constraints}
These prototype constraints demonstrate the flexibility and expressiveness of our constraint formalism.  The constraints below were chosen to highlight some common compliance measures and do not cover all possible constraints.  We encourage regulators to be creative when building constraints.

\bigskip
\noindent\textbf{Non-Membership Constraint (\texttt{NON-MEM})}: Proves that an address is not in a given set. For example, proving that the sender of this transaction is not on a list of sanctioned addresses.
$$C_{\text{non-mem}} = \text{sender}(\text{txn}) \notin p_{\text{sanctioned}}$$
where $p_{\text{sanctioned}}$ is the public parameter containing the sanction list.

\bigskip
\noindent\textbf{Membership Constraint (\texttt{MEM})}: Proves that an address is in a given set. For example, proving that the sender of this transaction is on an allow-list.
$$C_{\text{mem}} = \text{sender}(\text{txn}) \in p_{\text{allowlist}}$$
where $p_{\text{allowlist}}$ is the public parameter containing the allowed addresses.

\bigskip
\noindent\textbf{Protocol Interaction Constraint (\texttt{INTERACT})}: Proves that at some point in their transaction history, the sender of this transaction interacted with a specific protocol.
$$C_{\text{interact}} = \exists \, \text{t} \in \text{txn\_history}(\text{sender}(\text{txn}), [B_0, B_{\text{current}}]) : \text{recipient}(\text{t}) = p_{\text{protocol}}$$
where $p_{\text{protocol}}$ is the address of the required protocol.

\bigskip
\noindent\textbf{Protocol Avoidance Constraint (\texttt{AVOID})}: Proves that throughout their transaction history, the sender of this transaction never interacted with a specific protocol.
$$C_{\text{avoid}} = \forall \, \text{t} \in \text{txn\_history}(\text{sender}(\text{txn}), [B_0, B_{\text{current}}]) : \text{recipient}(\text{t}) \neq p_{\text{protocol}}$$
where $p_{\text{protocol}}$ is the address of the prohibited protocol.

\bigskip
\noindent\textbf{Age Constraint (\texttt{AGE})}: 
Requires that an address be at least a certain age. For example, require the sender's first transaction was at least 6 months ago as an effort to reduce sybil addresses.
$$C_{\text{age}} = \exists \, t \in \text{txn\_history}(\text{sender}(\text{txn}), [B_0, B_{\text{current}} - p_{\text{min-age}}]) : \text{sender}(t) = \text{sender}(\text{txn})$$
where $p_{\text{min-age}}$ is the minimum required age in blocks. The constraint is satisfied if any transaction from the sender exists before the age threshold, proving the address has been active for at least $p_{\text{min-age}}$ blocks.

\bigskip
\noindent\textbf{Structuring Constraint (\texttt{STRUCTURE})}: Specific example of US Bank Secrecy Act non-structuring requirement. Proves the sender of this transaction has not sent multiple transactions totaling over \$10,000 within a window of time to the same recipient.

\begin{align*}
C_{\text{structure}} = \forall \, r \in \text{recipients}(&s, [B_j, B_k] \subseteq [B_0, B_{\text{current}}] : (B_k - B_j) \leq p_{\text{window}}) :\\
&\sum_{\text{txn} \to r} \text{amount}(\text{txn}) < p_{\text{threshold}}
\end{align*}
where $i = \{[B_j, B_k] \subseteq [B_0, B_{\text{current}}] : (B_k - B_j) \leq p_{\text{window}}\}$ is a state set, $p_{\text{window}}$ is the time window (e.g., 24 hours in blocks), and $p_{\text{threshold}} = 10000$ USD equivalent. The state set triggers implicit universal quantification: the predicate must hold for all valid windows.

\subsection{Integrating applications}
In this prototype, applications integrate this framework by requiring proofs of compliance at specified entrypoint functions.  Applications choose compliance definitions based on the applications regulatory goals.

In the future, we envision a constraint compiler that compiles constraints from a formal syntax into custom circuits (see Future Work~\ref{sec:DSL}), but for prototype purposes, constraints are ``hand compiled" into Noir circuits.  Constraints are still listed to show the logical definition.

Each compliance definition $R$ is compiled as a ZK circuit using Noir and then published and maintained in a separate verifier Solidity contract, allowing the integrating application to reference compliance logic without implementing it directly.

\subsubsection{Example: Compliant Stablecoin}
The author of this stablecoin contract is, for example, a US based company and wishes to only allow users who are not on a list of US sanctioned addresses $p_{\text{sanctioned}}$ to interact with this stablecoin. 

The compliance definition requires both sender and recipient to satisfy the non-membership constraint:
$$R_{\text{stablecoin}} = C_{\text{non-mem}}(\text{sender}) \land C_{\text{non-mem}}(\text{recipient})$$
where both constraints evaluate against the same sanction list $p_{\text{sanctioned}}$.

A regulatory body deploys the $R_{\text{stablecoin}}\text{Verifier}$ contract as well as publishing the circuit to generate proofs of $R_{\text{stablecoin}}$.  The stablecoin contract author deploys the $\id{CompliantStablecoin}$ contract and sets $\id{verifierContract}$ to the address of $R_{\text{stablecoin}}\text{Verifier}$.  Users who with to transfer $\id{CompliantStablecoin}$ must first generate a proof of $R_{\text{stablecoin}}$ and supply that proof as an argument to the $\id{Transfer}$ function.

\bigskip
\begin{codebox}
\Procname{Contract: CompliantStablecoin}
\li \Comment State: 
\li $\id{balance}$ (mapping from address to uint)
\li $\id{verifierContract}$ (address of $R_{\text{stablecoin}}\text{ComplianceDefinition Verifier}$)
\li
\li \Comment User calls transfer with normal arguments and the compliance proof they generated
\li $\proc{Transfer}(\id{recipient}, \id{amount}, \id{proof}, \id{publicInputs})$
\li \>$\id{sender} \gets \id{tx.origin}$
\li \>
\li \>\Comment Verify compliance through external verifier contract
\li \>\kw{assert} $\id{verifierContract}.\proc{VerifyCompliance}(\id{sender}, \id{recipient}, \id{proof}, \id{publicInputs})$
\li \>
\li \>\Comment Execute transfer
\li \>\kw{assert} $\id{balance}[\id{sender}] \geq \id{amount}$
\li \>$\id{balance}[\id{sender}] \gets \id{balance}[\id{sender}] - \id{amount}$
\li \>$\id{balance}[\id{recipient}] \gets \id{balance}[\id{recipient}] + \id{amount}$
\li \>
\li \>\Return \const{true}
\end{codebox}

This architecture demonstrates the composability of the framework: the \texttt{ComplianceDefinition} contract $R_{\text{stablecoin}}$ is maintained independently from the stablecoin logic, allowing the sanction list to be updated without modifying the stablecoin contract itself.

\section{Benchmarks}
Proof generation in this framework relies on Noir's Barretenberg proving library, which supports two execution modes: a native binary backend and a WebAssembly (WASM) backend. These two modes execute the same cryptographic prover but in fundamentally different runtime environments, producing meaningfully different performance characteristics.  Our framework makes no requirements on whether proofs should be generated locally (native) or in-browser (WASM), so we present benchmarks for both options.
\begin{itemize}
    \item \textbf{Native}: The native backend runs Barretenberg as a platform-specific binary, leveraging native CPU instructions and OS-level threading. This represents the fastest possible proving performance on a given machine.
    \item \textbf{WASM}: The WASM backend runs Barretenberg as WebAssembly, which is how proofs are generated in browser-based applications. WASM execution incurs overhead from the sandboxed runtime, and browser environments introduce additional constraints: Web Worker threading coordination, competition with the DOM for resources, and tighter memory limits.
\end{itemize}

Both backends use identical circuit inputs, the same compiled circuit artifacts, and the same measurement methodology: timing witness generation  and proof generation as separate phases.  This controlled comparison isolates the performance impact of the runtime environment from the circuit complexity itself.

Barretenberg initialization time is the one-time cost of booting the proving runtime and is recorded separately. This cost is negligible for the native backend but averages 1.11 seconds for WASM, as it includes loading and compiling the WASM module. In a browser context, users pay this cost only on their first proof in a browser session.  This initialization time is not included in the below benchmark data.

\subsection{\texttt{MEM} constraint}

The circuit for the membership constraint is a single a Merkle proof of a 32 height tree.  

\begin{table}[h]
\centering
\begin{tabular}{l rrrr rrrr}
 & \multicolumn{4}{c}{\textbf{Native}} & \multicolumn{4}{c}{\textbf{WASM}} \\
\cmidrule(lr){2-5} \cmidrule(lr){6-9}
 & \textbf{Mean} & \textbf{Min} & \textbf{Max} & \textbf{Std Dev}
 & \textbf{Mean} & \textbf{Min} & \textbf{Max} & \textbf{Std Dev} \\
\midrule
Witness Generation & 0.01 & 0.00 & 0.03 & 0.01 & 0.01 & 0.01 & 0.01 & 0.00 \\
Proof Generation   & 0.20 & 0.18 & 0.27 & 0.02 & 0.39 & 0.35 & 0.45 & 0.03 \\
\midrule
\textbf{Total}     & \textbf{0.21} & 0.19 & 0.30 & 0.03 & \textbf{0.40} & 0.36 & 0.46 & 0.03 \\
\end{tabular}
\caption{Benchmark results for \texttt{MEM} constraint (times in seconds, $n=10$ runs).}
\label{tab:mem-nonmem-benchmarks}
\end{table}

\subsection{\texttt{NON-MEM} constraint}
Similarly, the circuit for the non-membership is 2 merkle proofs of a 32 height tree of adjacent leaves on a sorted merkle tree to prove an address does not exist between them.  We see it is takes roughly twice the time of \texttt{MEM} because it makes twice the merkle proofs.

\begin{table}[h]
\centering
\begin{tabular}{l rrrr rrrr}
 & \multicolumn{4}{c}{\textbf{Native}} & \multicolumn{4}{c}{\textbf{WASM}} \\
\cmidrule(lr){2-5} \cmidrule(lr){6-9}
 & \textbf{Mean} & \textbf{Min} & \textbf{Max} & \textbf{Std Dev}
 & \textbf{Mean} & \textbf{Min} & \textbf{Max} & \textbf{Std Dev} \\
\midrule
Witness Generation & 0.02 & 0.02 & 0.06 & 0.01 & 0.02 & 0.02 & 0.02 & 0.00 \\
Proof Generation   & 0.45 & 0.42 & 0.49 & 0.03 & 0.83 & 0.77 & 0.89 & 0.03 \\
\midrule
\textbf{Total}     & \textbf{0.48} & 0.44 & 0.55 & 0.03 & \textbf{0.85} & 0.79 & 0.91 & 0.03 \\
\end{tabular}
\caption{Benchmark results for \texttt{NON-MEM} constraint (times in seconds, $n=10$ runs).}
\label{tab:mem-nonmem-benchmarks}
\end{table}

\section{Future Work}

Our framework is a start on creating a means to bring
compliance and privacy together in a way that meets perceived social needs for both.  The components of the framework will benefit from further work addressing both efficiency and breadth of coverage.

\subsection{Constraint Domain-Specific Language}
\label{sec:DSL}
To transform written law into a zkSNARK there needs to be a human readable intermediary syntax.  The constraint system introduced in \ref{sec:constraints} can be extended into a novel DSL used by regulatory bodies to define compliance.  A goal of this DSL is to be easier to read than the formal expression of constraints in raw logic.  This would make constraint construction \textbf{much} easier for regulators, leading to more framework integration.  A larger ecosystem of compliance definitions and a readable compliance syntax would in turn give applications and users more confidence.

This DSL would compile directly into constraints.  The addition of this proposed DSL would require no changes to the overall framework and is backward compatible: the framework already is built around the constraint unit.  Even the verifier contracts could stay the same.  The only modification that would be necessary is a patch to the proof manager to add this compilation step.

\reminder{TODO: highlight how we're interoperable with any ZK backend.  Could target RISCV as the output of our compiler to feed into any RISCV zkvm}

\paragraph{Example \texttt{NON-MEM} constraint translation} 

The non-membership constraint defined as:
$$C_{\text{non-mem}} = \text{sender}(\text{txn}) \notin p_{\text{sanctioned}}$$
Could instead be expressed in a constraint DSL in a more readable fashion:

$$C_{\text{non-mem}} = \mathtt{User\ is\ not\ in\ } p_{\text{sanctioned}}$$

Where ``is" expresses the temporal context of the current block, and $p_{\text{sanctioned}}$ is the sanction list.

\paragraph{Example \texttt{INTERACT} constraint translation} 

The protocol interaction constraint defined as:

$$C_{\text{interact}} = \exists \, \text{t} \in \text{txn\_history}(\text{sender}(\text{txn}), [B_0, B_{\text{current}}]) : \text{recipient}(\text{t}) = p_{\text{protocol}}$$

Could be translated into:
$$C_{\text{interact}} = \mathtt{User\ has\ sent\ transaction\ to\ } p_{\text{protocol}}$$

Where ``has" means at any point in the user's transaction history and $p_{\text{protocol}}$ is the address of the required protocol.

\paragraph{Example published compliance definition}

Compare the compliance definition $R = C_\text{non-mem} \land C_\text{interact}$ with the current constraint syntax to the same definition using a constraint DSL.  

Which is easier to understand?

\begin{codebox}
\Procname{Current:}
\zi \centering $R = \left(\text{sender}(\text{txn}) \notin p_{\text{sanctioned}}\right) \;\land\; \left(\exists \, \text{t} \in \text{txn\_history}(\text{sender}(\text{txn}), [B_0, B_{\text{current}}]) : \text{recipient}(\text{t}) = p_{\text{protocol}}\right)$
\end{codebox}

\begin{codebox}
\Procname{Using constraint DSL:}
\zi \centering $R = \left(\mathtt{User\ is\ not\ in\ } p_{\text{sanctioned}}\right) \;\mathtt{AND}\; \left(\mathtt{User\ has\ sent\ transaction\ to\ } p_{\text{protocol}}\right)$
\end{codebox}

\subsection{Proof inference and optimization}
If several constraints need to be proved, one can just prove their conjunction, but there is likely benefit in transforming a given conjunction of constraints into a shorter and more easily proved form.  A similar issue arises with updates to constraints.
Say a regulator's compliance definition $R_1$ is updated to $R_2$.  They might not only publish the full $R_2$ for new proofs but also $R_{1,2}$ that has the property that $R_1\ \land\ R_{1,2} \implies R_2$.  Assuming $R_{1,2}$ is less costly to prove than $R_2$, a user who previously proved $R_1$ will be rewarded with faster passing of $R_2$ via the logical implication.  
Similarly, if certain facts about a user's transaction history have already been proved, those proofs can be reused.
These considerations, at scale, create a need for an 
indexing of proofs based on what has been proven about what user or transaction history the proof pertains.  Such indexing is much more than a simple database-style lookup.

For complex constraint systems, a transaction's proof obligation may consume a significant fraction of
overall transaction latency.  For such complex
constraint systems, any optimization algorithm must take into account not only simplification of statements
in boolean logic, but also the estimated cost to generate each component proof.  A simpler logic expression might be over a dataset leading to a larger circuit than a less-simple logic expression.  The resulting optimization program has features closely related to database query optimization, though with a different cost model.  Future work will adapt known optimization techniques from database systems \cite{DSC} to this novel optimization domain.  

Also related to the optimization problem is the underlying execution system.  We generate proofs by
submitting code in the ZK-friendly domain-specific language Noir \cite{AztecNoir22}. After generating an efficient high-level proof strategy, the actual proofs needed to be coded into Noir in a manner ensuring the fastest possible proof generation by the underlying prover.  Aztec, Noir's developer, has published documents on Noir optimization (\url{https://noir-lang.org/docs/explainers/explainer-writing-noir}), but we anticipate further work to generate optimal Noir code for our framework.

\subsection{Backend provers}
The framework is designed independent of any ZK prover.  By targeting the Noir ZK-DSL, any PLONK-based prover or R1CS-based prover can be used directly.  Future work includes broadening our set of provers further, taking into account issues of efficiency and target chains.

\subsection{Carry compliance proof through mixers}
Commitment tree mixers like Tornado Cash~\cite{tornadocash} and Privacy Pools~\cite{buterin2023blockchain} unlink the sender and receiver addresses of a transfer, creating privacy for the user.  A user enters the mixer by sending funds to the mixer along with a secret.  An address can withdraw funds from the mixer if they provide a ZK proof that know a secret corresponding to a deposit.  There is no way to reveal association between any two deposit and withdraw transactions in a mixer without the user revealing their secret.  The common use case is a privacy-desiring user withdrawing to an addresses with zero transaction history to obfuscate the association between the two addresses.

Inspired by ``Derecho: Privacy Pools with Proof-Carrying Disclosures"~\cite{10.1145/3658644.3670270}, the compliance properties required to enter the mixer can be transferred to the withdrawing address by proving the owner of the withdrawing address knows the secret of a leaf in the commitment tree, therefore proving that they have access to an account that met the compliance definition to deposit into the mixer.  Then an additional proof needs to be made to assure that the withdrawing account hasn't previously violated compliance constraints.  Only the constraints required for depositing to the mixer can be carried to the withdrawing address.

\textbf{Example}: Alice signs up to Coinbase and provides required KYC documents.  Alice then receives funds at her on-chain address from a known Coinbase address.  Alice can trivially prove her address received funds from Coinbase.  Alice does not want her entire transaction history to be public so she deposits some funds in a mixer that has integrated this framework and requires a compliance definition with a single constraint: $C_{\text{origin}}$

$$C_{\text{origin}} = \exists \, t \in \text{first-txn}(\text{sender}(\text{txn}), [B_0, B_{\text{current}}]) : \text{sender}(t) \in A$$

$C_{\text{origin}}$ is read as ``the sender of this transaction's first transaction was from a set of addresses $A$'', where $A$ is a set of public Coinbase addresses.

Alice proves this to deposit funds to the mixer.  She later withdraws her funds from the mixer to a freshly generated address with no transaction history.  Even though this new address never received funds from Coinbase, Alice is able to supply her commitment tree secret into a circuit to prove that she is transferring from an address that successfully proved $C_{\text{origin}}$, and therefore should be able to prove $C_{\text{origin}}$ when interacting with protocols with her new address.  Work needs to be done testing and implementing this.

\subsection{Alternative Applications of Constraint Proving}
This paper focused on the use case of proving compliance over a transaction history.  However, the underlying mechanism of proving arbitrary properties of a transaction history is powerful and can be used in other scenarios.  For example, airdrops that only allow addresses with specific properties to claim.  It could potentially even move some smart-contract execution off-chain by requiring a proof that an invariant holds for a particular address instead of calculating it on-chain.

\subsection{Latency Optimization}
Several of the topics in our future work agenda relate to performance.  Since proofs in our framework are generated as part of the full execution of a transaction, latency as measured from time of transaction submission to the achievement of a user-acceptable level of finality, is extremely important.  Merging the above agendas into a full end-to-end optimization and benchmarking strategy is highly important.

\subsection{Industrial and Regulator Acceptance}
The ultimate success of our framework rests on its acceptance by regulators, industrial application developers, and users.  We have had positive preliminary discussions, but expect that the strongest
driver for acceptance will be the competitive ecosystem.  Our framework enables stablecoin systems
to complete in offering privacy with transparent proof of legality.  Regulatory domains seeking to attract financial activity can use our framework to facilitate compliance both in terms of system-development costs and in terms of the cost of compliance verification.

\bibliographystyle{abbrv}
\bibliography{blockchain}
\end{document}

